### 2015-11-30 paper/paper.Rmd
start

---
title: R Packages to Aid in Handling Web Access Logs
author:
  - name: Oliver Keyes
    affiliation: Wikimedia Foundation
    address:
    - line 1
    - line 2
    email:  author1@work
  - name: Bob Rudis
    affiliation: Rapid7
    address:
    - line 1
    - line 2
    email:  author2@work
abstract: >
  Web access logs contain information on HTTP(S) requests and form a key part of both industry and academic explorations
  of human behaviour on the internet - explorations commonly performed in R. In this paper we explain and demonstrate
  a series of packages - **webreadr**, **urltools**, **iptools** and **rgeolocate** - designed to efficiently read in, parse
  and munge access log data, allowing researchers to handle it easily.
preamble: >
output: rticles::rjournal_article
---

## Introduction

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

## Reading access logs

This section may contain a figure such as Figure \ref{figure:rlogo}.

\begin{figure}[htbp]
  \centering
  \includegraphics{Rlogo}
  \caption{The logo of R.}
  \label{figure:rlogo}
\end{figure}

## Decoding and parsing URLs

There will likely be several sections, perhaps including code snippets, such as:

```{r}
x <- 1:10
x
```

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](http://journal.r-project.org/latex/RJauthorguide.pdf).

\bibliography{RJreferences}