### 2015-12-05 paper/paper.tex
add

% !TeX root = RJwrapper.tex
\title{R Packages to Aid in Handling Web Access Logs}
\author{by Oliver Keyes, Bob Rudis}

\maketitle

\abstract{
Web access logs contain information on HTTP(S) requests and form a key
part of both industry and academic explorations of human behaviour on
the internet - explorations commonly performed in R. In this paper we
explain and demonstrate a series of packages - \textbf{webreadr},
\textbf{urltools}, \textbf{iptools} and \textbf{rgeolocate} - designed
to efficiently read in, parse and munge access log data, allowing
researchers to handle it easily.
}

\subsection{Introduction}\label{introduction}

The rise of the World Wide Web over the last few decades has made it
dramatically easier to access and transfer data, and R \citep{RCore}
boasts abundant functionality when it comes to taking data \emph{from}
the web. Base R itself has simple file downloading and page reading
capabilities, through the \texttt{download.file} and \texttt{readLines}
functions, and additional functionality is made available for handling
web-accessible data through packages such as \textbf{httr} \citep{httr}.

Data \emph{on} the web is not, however, the only kind of web data that
interests researchers; web traffic is, in and of itself, an interesting
data source. Access logs - records of connections between users and a
web server - are an asset and resource for people studying everything
from user behaviour on the internet \citep{halfak}, to website
performance \citep{performance}, to information security
\citep{infosec}.

As a statistically-oriented programming language, R is commonly used by
these same researchers for data analysis, testing and reporting - but it
lacks tools designed for the kinds of data sources and formats that are
encountered when dealing with access logs. In this article we review the
use cases for particular operations over web data, the limitations in
base R when it comes to performing those operations, and a suite of R
packages designed to overcome them: \textbf{webreadr} \citep{webreadr},
designed for reading access logs in, \textbf{urltools} \citep{urltools},
for URL manipulation, \textbf{iptools} \citep{rgeolocate} for handling
IP addresses, and \textbf{rgeolocate} \citep{rgeolocate} for direct IP
geolocation.

\subsection{Reading access logs}\label{reading-access-logs}

The first task with any data analysis is to read the data into R so that
it can be manipulated. With access logs this is slightly complicated by
the fact that there is no one standard for what a log should look like;
instead, there are multiple competing approaches from different software
platforms and eras. These include the Common Log Format (CLF), the
confusingly-named Combined Log Format, and formats used by individual,
commonly-used software platforms - such as the custom format for the
Squid internet caching software, and the format used by Amazon Web
Services (AWS).

The difference between formats can easily be shown by looking at how
timestamps are represented:

\begin{table}[ht]
\centering
\caption{Timestamps in Common Access Log Formats}
\begin{tabular}{rlr}
  \hline
Log Type & Timestamp Columns & Timestamp Format \\ 
  \hline
Common Log Format & 1 & 10/Oct/2000:13:55:36 -0700\\ 
Combined Log Format & 1 & 26/Apr/2000:00:23:48 -0400\\ 
Squid & 1 & 1286536309.450\\ 
AWS & 2 & 2014-05-23    01:13:11\\ 
   \hline
\end{tabular}
\end{table}

With four log types, we have three different timestamp formats - and
that's only one of the many columns that could appear. These logs also
vary in whether they specify quoting fields (or sanitising unquoted
ones), the columns they contain and the data each column contains in
turn.

To make reading access logs into R as easy as possible we created the
\textbf{webreadr} package. This contains user-friendly equivalents to
\texttt{read.table} for each type of log, detecting the fields that
should appear, converting the timestamps into POSIX objects, and merging
fields or splitting fields where necessary. The package contains four
core functions, one for each form of log: \texttt{read\_clf} for the
Common Log Format, \texttt{read\_combined} for the Combined Log Format,
and \texttt{read\_squid} and \texttt{read\_aws} for Squid and AWS
formats respectively. Each one abstracts away the complexity of
specifying column names and formats, and instead allows a researcher to
read a file in with a minimal amount of work.

As the name suggests, it is built not on top of base R but on top of the
\textbf{readr} \citep{readr} package, allowing us to take advantage of
substantial speed improvements that package's base functions have over
base R. These improvements can be seen in the visualisation below, which
uses \textbf{microbenchmark} \citep{microbenchmark} to compare 100 reads
of a 600,000-line ``squid'' formatted file with webreadr to the same
operation performed in base R:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{reading_benchmarks}
    \caption{Results of microbenchmark run: read\char`_squid versus base-R equivalent code}
\end{figure}

\textbf{webreadr} takes 1.2 seconds to read this file in at a minimum,
and 1.48 at maximum, with a median of 1.33. This is consistently 3.5-6
times faster than the equivalent base R functionality - and is also, as
explained, far easier to use.

\subsection{Decoding and parsing URLs}\label{decoding-and-parsing-urls}

Within access logs, URLs appear - both to describe the web asset or page
that the user requested and where the user came from. These fields are
usually named url and referer respectively.

\subsubsection{Decoding}\label{decoding}

Both values can be percent-encoded, allowing them toinclude characters
that are valid but reserved by the URL specification as having special
meanings (``reserved characters''). A \texttt{\#} symbol, for example,
is encoded as \%23: a percentage symbol, followed by the byte-code for
that character.

The encoding of reserved characters is useful, since it means that URL
paths and queries can contain a vast array of values - but it makes data
analysis tougher to do. Examples of common data analysis or cleaning
operations that become more difficult are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  \textbf{Aggregation}. Aggregating URLs together is useful to identify,
  for example, the relative usage and popularity of particular pages in
  your data - but it becomes tougher if encoding is inconsistent,
  because two URLs could hold the same value but \emph{look} very
  different.
\item
  \textbf{Value selection}. With text-based data, regular expressions
  are a common way of filtering or selecting entries that meet
  particular conditions, but things become fuzzy when you have to look
  not just for particular characters (a space, say) but also the encoded
  values (\%20).
\item
  \textbf{Exploratory data analysis} (EDA). EDA is a common initial step
  to investigate a data set, examining the variables and values it
  contains and whether they meet a researcher's expectations - but on a
  practical basis it becomes difficult when the values aren't
  human-readable.
\end{enumerate}

The solution is to be able to consistently decode URLs, which makes
URL-based data far easier to analyse. Base R does contain a function,
URLdecode, for doing just that, but is neither vectorised nor based on
compiled code; with large datasets it takes an extremely long time.

To solve this common problem in analysing request logs, the
\textbf{urltools} {[}@urltools{]} package was created. This contains a
function, url\_decode, which decodes URLs - but relies on vectorised,
compiled code. Benchmarking the two approaches against each other shows
that the \texttt{urltools} implementation is approximately 60-70 times
faster over large datasets. Again using \textbf{microbenchmark}, if we
compare the vectorised decoding of 1,000,000 URLs with \texttt{urltools}
against \texttt{URLdecode} in a \texttt{vapply} loop, we see a 60-70
times speed improvement of urltools over base R:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{decoding_benchmarks}
    \caption{Results of microbenchmark run: url\char`_decode versus base-R equivalent code}
\end{figure}

\subsubsection{Parsing}\label{parsing}

The standard for URLs \citep{RFC1738} divides them into a heirarchical
sequence of components - the scheme (``http''), host
(``en.wikipedia.org''), port (``800''), path (``wiki/Main\_Page'') and
searchpart, or parameters (``action=edit''). Together, these make up a
URL (``\url{http://en.wikipedia.org:800/wiki/Main_Page?action=edit}'').

Parsing URLs to isolate and extract these components is a useful ability
when it comes to exploring request logs; it lets a researcher pick out
particular schemes, paths, hosts or other components to aggregate by,
identifying how users are behaving and what they are visiting. It makes
anonymising data - by removing, for example, the parameters or path,
which can contain relatively unique information - easier.

Base R does not have native code to parse URLs, but the \textbf{httr}
package \citep{httr} contains a function, \texttt{parse\_url}, designed
to do just that and built on base R's capabilities. This function is
neither vectorised nor based on compiled code, and produces a list,
making looping over a set of URLs to parse each one a somewhat thorny
experience. This is totally understandable given the intent behind that
function, which is to decompose individual URLs within the context of
making HTTP requests, rather than to analyse request logs \emph{en
masse}.

\texttt{urltools} contains \texttt{url\_parse} - which does the same
thing as the equivalent \texttt{httr} functionality, but in a vectorised
way, relying on compiled code, and producing a data.frame. Within the
context of parsing and processing access logs, this is far more useful,
because it works efficiently over large sets: httr's functionality, not
having designed to be, does not. Indeed, benchmarking showed that httr's
version is (for large vectors, at least) approximately 570 times slower
than \texttt{url\_parse}:

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.4]{decoding_benchmarks}
    \caption{Results of microbenchmark run: url\char`_parse versus httr equivalent code}
\end{figure}

\subsection{Summary}\label{summary}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions
for Authors}.

\bibliography{RJreferences}

\address{
Oliver Keyes\\
Wikimedia Foundation\\
line 1\\ line 2\\
}
\href{mailto:author1@work}{\nolinkurl{author1@work}}

\address{
Bob Rudis\\
Rapid7\\
line 1\\ line 2\\
}
\href{mailto:author2@work}{\nolinkurl{author2@work}}